---
title: "Penguins Project"
author: "Jackson Small"
date: "2023-11-28"
output:
  word_document: default
  pdf_document: default
---

Removing Environment Variables (optional)
```{r}
rm(list=ls())
```


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```


Needed Libraies 
```{r}
library(ggplot2)
library(GGally)
library(car)

```

```{r}
penguins <- read.csv(file="project2/penguins.csv", header = TRUE)

head(penguins,10)

```


Frequency Exploration

```{r}

summary(penguins)
```

```{r}
summary(penguins$bill_length_mm) # summary chart for bill length
summary(penguins$bill_depth_mm) # summary chart for bill depth
summary(penguins$flipper_length_mm) # summary chart for flipper length
summary(penguins$body_mass_g) # summary chart for body mass (g)

```


General Graphs
```{r}
library(ggplot2)

body_mass_plot <- ggplot(data = penguins) + # distribution of the response variable
  aes(x = (body_mass_g)) +
  geom_histogram(color = "red", binwidth = 50) +
  xlab("Body Mass (g)")
body_mass_plot

bill_length_plot <- ggplot(data = penguins) + # distribution of bill length variable
  aes(x = bill_length_mm) +
  geom_histogram(color = "blue", binwidth = 0.5) +
  xlab("Bill Length (mm)")
bill_length_plot

bill_depth_plot <- ggplot(data = penguins) + # distribution of bill depth 
  aes(x = bill_depth_mm) +
  geom_histogram(color = "yellow", binwidth = 0.5) +
  xlab("Bill depth (mm)")
bill_depth_plot

flipper_length_plot <- ggplot(data = penguins) + # distribution of flipper length
  aes(x = flipper_length_mm) +
  geom_histogram(color = "purple", binwidth = 0.5) +
  xlab("Flipper Length (mm)")
flipper_length_plot





```

Scatter Plots

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(title = "Bill Length vs. Body Mass",
       x = "Bill Length (mm)",
       y = "Body Mass (g)") +
  theme_linedraw()

ggplot(penguins, aes(x = bill_depth_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(title = "Bill Depth vs. Body Mass",
       x = "Bill Depth (mm)",
       y = "Body Mass (g)") +
  theme_linedraw()

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs. Body Mass",
       x = "Bill Depth (mm)",
       y = "Body Mass (g)") +
  theme_linedraw()
```

Pair Plot
```{r}
library(GGally) 
ggcorr(penguins, method = c("pairwise"))
ggpairs(data = penguins,
        columns = c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g"),
        mapping = aes(colour = species)) +
  theme_minimal()  
```


Exploring General Assumptions

```{r}
lm_full <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm,
               data = penguins)

plot(lm_full)
```

The Five Statistical Assumptions
Existence - For every x value there is a y value which creates a finite mean and variance
Independence - met 
Linearity - met
Homoscedasticity - met
Normality - met with histograms

Collinearity Evaluation 
```{r}
vif(lm_full) 
# all values are < 10 therefore there are no collinearity risks
```


Data Preparation: 
```{r}
colSums(is.na(penguins))

```

```{r}
# From the plot(lm_full) we can see that observation 294 has high leverage so lets remove it and all NA's

penguins_no294 <- penguins[-294, ]

penguins_clean <- na.omit(penguins_no294)
head(penguins_clean, 10)

lm_full_clean <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm,
               data = penguins_clean)

plot(lm_full_clean)

summary(lm_full_clean)
summary(penguins_clean)


# to fix the bend in the residuals vs fitted graph. As we can see the difference is shown

```






Univariate Models!
```{r}
lm_bill_length <- lm(body_mass_g ~ bill_length_mm, data = penguins_clean)
summary(lm_bill_length)
```

```{r}
lm_bill_depth <- lm(body_mass_g ~ bill_depth_mm, data = penguins_clean)
summary(lm_bill_depth)
```

```{r}
lm_flipper_length <- lm(body_mass_g ~ flipper_length_mm, data = penguins_clean)
summary(lm_flipper_length)
```


Full Models!
```{r}
lm_full_clean <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = penguins_clean)

summary(penguins_clean)
summary(lm_full_clean)
anova(lm_full_clean)

```

interaction testing
```{r}
lm_interaction <- lm(body_mass_g ~ bill_length_mm + bill_depth_mm + flipper_length_mm + (bill_length_mm * bill_depth_mm), data = penguins_clean)
summary(lm_interaction)

BWDfit.p <- ols_step_backward_p(lm_interaction, prem = 0.05)
BWDfit.p

BWDfit.p <- ols_step_backward_p(lm_interaction, prem = 0.05, details = TRUE)

```



From looking at the Univariate models it shows that Bill length, Bill depth, and flipper length are all linear. When we create the model (y) = b0 + b1(bill_length) +b2(bill_depth) + b3(flipper_length).

stepwise elimination:

```{r}
list_lm <- list()
list_lm[["0"]] <- lm_full_step <- lm(
  formula = body_mass_g~ 1,
  data = penguins_clean
)
list_lm[["1"]] <- lm_full_step <- lm(
  formula = body_mass_g~ bill_length_mm,
  data = penguins_clean
)
list_lm[["2"]] <- lm_full_step <- lm(
  formula = body_mass_g~ bill_length_mm + bill_depth_mm,
  data = penguins_clean
)

list_lm[["3"]] <- lm_full_step <- lm(
  formula = body_mass_g~ bill_length_mm + bill_depth_mm + flipper_length_mm,
  data = penguins_clean
)



lapply(
  X = list_lm,
  FUN = summary
)


sapply(
  X = list_lm,
  FUN = extractAIC,
  k = 2
)
```



backwards elimination: 


```{r}
library(olsrr)
BWDfit.p <- ols_step_backward_p(lm_full_clean, prem = 0.05)
BWDfit.p

BWDfit.p <- ols_step_backward_p(lm_full_clean, prem = 0.05, details = TRUE)

```
Overall we can conclude that the best model and final model is 

```{r}
lm_final <- lm(body_mass_g ~ flipper_length_mm, data = penguins_clean)
summary(lm_final)
anova(lm_final)
```


Boxplots by Species 
```{r}

boxplot(body_mass_g ~ species,
        data = penguins_clean, 
        main="Body Mass (g) by Species",
        xlab="Species", 
        ylab="Body Mass (g)",
        col = c('red', 'gold', 'orange'))

```


